---
title: "R Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(lme4)
library(effects)
library(corrplot)
library(knitr)
library(ggplot2)
#install.packages('pROC')
library(pROC)
library(MASS)
library(class)
#library(klaR)
#install.packages('earth')
library(earth)
```

```{r}

Pokemon=read.csv("Pokemon.csv")
levels(Pokemon$type2)=c(levels(Pokemon$type2),"none")
Pokemon$type2[Pokemon$type2==""]="none"
#summary(Pokemon)
#Pokemon$name[is.na(Pokemon$weight_kg)]
#pairs(Pokemon[,34:ncol(Pokemon)])
Pokemon_Cor=Pokemon[,c("base_total","weight_kg","height_m","base_happiness","base_egg_steps","experience_growth","Evolution","capture_rate")]
Pokemon_Cor$generation=as.numeric(Pokemon$generation)
Pokemon_Cor$generation=as.numeric(Pokemon$is_legendary)
Pokemon_Cor$capture_rate=as.integer(Pokemon$capture_rate)
corrplot(cor(Pokemon_Cor))

means<-dplyr::select(Pokemon,c("type1","base_total")) %>%
group_by(.,type1)%>%
summarise(.,mean(base_total))

op <- par(no.readonly = TRUE)
par(mar=c(8, 4, 2, 2) + 0.1)
plot(means$type1,means$`mean(base_total)`,xlab="type",xaxt="n",ylab="total",main="Total Stats by Type 1")
axis(1, at=1:nrow(means), labels=means$type1, las = 2, cex.axis = 0.8)
#reset settings
par(op)

means<-dplyr::select(Pokemon,c("type2","base_total")) %>%
group_by(.,type2)%>%
summarise(.,mean(base_total))

op <- par(no.readonly = TRUE)
par(mar=c(8, 4, 2, 2) + 0.1)
plot(means$type2,means$`mean(base_total)`,xlab="type",xaxt="n",ylab="total",main="Total Stats by Type 2")
axis(1, at=1:nrow(means)+1, labels=means$type2, las = 2, cex.axis = 0.8)
#reset settings
par(op)

Pokemon$capture_rate=scale(as.integer(Pokemon$capture_rate))
#Pokemon$type2[is.na(Pokemon$type2)]="none"
Pokemon$generation=as.factor(Pokemon$generation)
Pokemon$is_legendary=as.factor(Pokemon$is_legendary)

#pairs(Pokemon[,34:ncol(Pokemon)])
pairs(Pokemon[,c("base_total","weight_kg","height_m",  "percentage_male","base_happiness","base_egg_steps","experience_growth")])
Pokemon$percentage_male[is.na(Pokemon$percentage_male)]=0
Pokemon$base_happiness=c(scale(Pokemon$base_happiness))
Pokemon$base_egg_steps=c(scale(Pokemon$base_egg_steps))
Pokemon$experience_growth=c(scale(Pokemon$experience_growth))

Pokemon$weight_kg=c(scale(Pokemon$weight_kg))
Pokemon$height_m=c(scale(Pokemon$height_m))
#pairs(Pokemon[,34:ncol(Pokemon)])
pairs(Pokemon[,c("base_total","weight_kg","height_m","base_happiness","base_egg_steps","experience_growth","Evolution","capture_rate")])
#Pokemon$base_happiness=1/(Pokemon$base_happiness)
#Pokemon$base_egg_steps=1/(Pokemon$base_egg_steps)
#Pokemon$experience_growth=1/(Pokemon$experience_growth)
#Pokemon$weight_kg=1/(Pokemon$weight_kg)
#Pokemon$height_m=1/(Pokemon$height_m)




Num_Col=c(26:29,32,34,35,36,37,39:41)
Pokemon_Cor=Pokemon[,c("base_total","weight_kg","height_m","base_happiness","base_egg_steps","experience_growth","Evolution","capture_rate")]
Pokemon_Cor$type1=as.numeric(Pokemon$type1)


corrplot(cor(Pokemon_Cor))

Pokemon_Cor2=Pokemon[,c("base_total","attack","defense","speed","hp","sp_attack","sp_defense")]
corrplot(cor(Pokemon_Cor2))
#summary(Pokemon_Cor)

```

```{r}

#Pokemon=read.csv("Pokemon.csv")
#pairs(Pokemon[,34:ncol(Pokemon)])
#Pokemon$capture_rate=scale(as.integer(Pokemon$capture_rate))
#Pokemon$type2[is.na(Pokemon$type2)]="none"
#levels(Pokemon$type2)=c(levels(Pokemon$type2),"none")
#Pokemon$type2[Pokemon$type2==""]="none"
#Pokemon$generation=as.factor(Pokemon$generation)

unique(Pokemon$percentage_male)
```

```{r}
#https://bradleyboehmke.github.io/HOML/mars.html
#Pokemon=Pokemon[,colnames(Pokemon)!="sp_attack"]
#Pokemon=Pokemon[,colnames(Pokemon)!="sp_defense"]
#Pokemon=Pokemon[,colnames(Pokemon)!="attack"]
#Pokemon=Pokemon[,colnames(Pokemon)!="hp"]
#Pokemon=Pokemon[,colnames(Pokemon)!="defense"]
#Pokemon=Pokemon[,colnames(Pokemon)!="base_total"]
#Pokemon=Pokemon[,colnames(Pokemon)!="abilities"]
#Pokemon=Pokemon[,colnames(Pokemon)!="base_total"]
#Pokemon=Pokemon[,colnames(Pokemon)!="classfication"]
#Pokemon=Pokemon[,colnames(Pokemon)!="japanese_name"]
#Pokemon=Pokemon[,colnames(Pokemon)!="name"]
#Pokemon=Pokemon[,colnames(Pokemon)!="pokedex_number"]
#Pokemon$perAtt=Pokemon$attack/Pokemon$base_total
#model=earth(defense~is_legendary+generation+percentage_male+weight_kg+height_m+experience_growth+speed,Pokemon[complete.cases(Pokemon),])
set.seed(801)
model_base_only=earth(base_total~type1*type2+height_m*weight_kg+is_legendary*experience_growth+generation+base_egg_steps*capture_rate+base_happiness+capture_rate+Evolution,Pokemon, nfold=10, ncross=50, varmod.method="lm")
Predict_model_base_only=predict(model_base_only,interval="pint")

model_toy_height=earth(defense~height_m+weight_kg,Pokemon[complete.cases(Pokemon),-1], nfold=10, ncross=50, varmod.method="lm")
Predict_model_toy_height=predict(model_toy_height,interval="pint")
plotmo(model_toy_height)
summary(model_toy_height)
plot(model_toy_height)
#model=step(lm(base_total~type1*type2+height_m*weight_kg+is_legendary*experience_growth+generation+base_egg_steps+base_happiness+capture_rate,Pokemon[complete.cases(Pokemon),-1]))
#model=earth(perAtt~type1*type2+height_m*weight_kg+is_legendary,Pokemon[complete.cases(Pokemon),-1])
#expand.bpairs(type1~defense,Pokemon[complete.cases(Pokemon),])
#model=earth(attack~.^2,Pokemon[complete.cases(Pokemon),-1])
summary(model_base_only)
plotmo(model_base_only)
plot(model_base_only)



Pokemon_Complete$Pred_Base_only<-Predict_model_base_only$fit
#View(Pokemon[complete.cases(Pokemon),])
```

Reasonable  R^2 when calculating total, bad when stratified by exact stat.  We could treat it as a missing data problem and predict one at a time based on best stat.

```{r}

Pokemon_Complete=Pokemon
Pokemon_Complete$Pred_base=Predict_model_base_only$fit
Predict_model_base_order=Predict_model_base_only[order(Pokemon_Complete$Pred_base),]
Pokemon_Complete_order=Pokemon_Complete[order(Pokemon_Complete$Pred_base),]
Pokemon_Complete_order$index=seq(1:nrow(Pokemon_Complete_order))

ggplot(Pokemon_Complete_order, aes(index, Pred_base))+
    geom_point()+
    geom_point(aes(y = base_total, colour=ifelse(base_total<Predict_model_base_order$lwr |
                                            base_total>Predict_model_base_order$upr,"orange","red")))+
    geom_line(data=Pokemon_Complete_order)+
    geom_ribbon(aes(ymin=Predict_model_base_order$lwr,ymax=Predict_model_base_order$upr),alpha=0.3) +xlab("Pokemon sorted by Predicted total")+
  ylab("Total Stats") + ggtitle("Predictions and True Values") + theme(legend.position = "none")+
  theme(plot.title = element_text(hjust = 0.5))

Pokemon_Complete_order$name[Pokemon_Complete_order$base_total>Predict_model_base_order$upr]
Pokemon_Complete_order$name[Pokemon_Complete_order$base_total<Predict_model_base_order$lwr]
Pokemon_Complete_order$name[Pokemon_Complete_order$base_total==max(Pokemon_Complete_order$base_total)]
Pokemon_Complete_order$name[Pokemon_Complete_order$Pred_base==min(Pokemon_Complete_order$Pred_base)]
Pokemon_Complete_order$base_total[Pokemon_Complete_order$name=="Rayquaza"]
Pokemon_Complete_order$Pred_base[Pokemon_Complete_order$name=="Rayquaza"]
```


```{r}
#With base
model_attack=earth(attack~type1*type2+height_m*weight_kg+is_legendary*experience_growth+generation+base_egg_steps*capture_rate+base_happiness+capture_rate+base_total,Pokemon[complete.cases(Pokemon),-1])
sum_mod=summary(model_attack)
paste("attack: ",sum_mod$rsq)

model_hp=earth(hp~type1*type2+height_m*weight_kg+is_legendary*experience_growth+generation+base_egg_steps*capture_rate+base_happiness+capture_rate+base_total,Pokemon[complete.cases(Pokemon),-1])
sum_mod=summary(model_hp)
paste("hp: ",sum_mod$rsq)

model_defense=earth(defense~type1*type2+height_m*weight_kg+is_legendary*experience_growth+generation+base_egg_steps*capture_rate+base_happiness+capture_rate+base_total,Pokemon[complete.cases(Pokemon),-1])
sum_mod=summary(model_defense)
paste("defense: ",sum_mod$rsq)

model_sp_attack=earth(sp_attack~type1*type2+height_m*weight_kg+is_legendary*experience_growth+generation+base_egg_steps*capture_rate+base_happiness+capture_rate+base_total,Pokemon[complete.cases(Pokemon),-1])
sum_mod=summary(model_sp_attack)
paste("sp_attack: ",sum_mod$rsq)

model_sp_defense=earth(sp_defense~type1*type2+height_m*weight_kg+is_legendary*experience_growth+generation+base_egg_steps*capture_rate+base_happiness+capture_rate+base_total,Pokemon[complete.cases(Pokemon),-1])
sum_mod=summary(model_sp_defense)
paste("sp_defense: ",sum_mod$rsq)

model_speed=earth(speed~type1*type2+height_m*weight_kg+is_legendary*experience_growth+generation+base_egg_steps*capture_rate+base_happiness+capture_rate+base_total,Pokemon[complete.cases(Pokemon),-1])
sum_mod=summary(model_speed)
paste("speed: ",sum_mod$rsq)
```

```{r}
set.seed(801)
#Without base
model_attack=earth(attack~type1*type2+height_m*weight_kg+is_legendary*experience_growth+generation+base_egg_steps*capture_rate+base_happiness+capture_rate+Evolution,Pokemon[complete.cases(Pokemon),-1], nfold=10, ncross=50, varmod.method="lm")
sum_mod=summary(model_attack)
paste("attack: ",sum_mod$rsq)

model_hp=earth(hp~type1*type2+height_m*weight_kg+is_legendary*experience_growth+generation+base_egg_steps*capture_rate+base_happiness+capture_rate+Evolution,Pokemon[complete.cases(Pokemon),-1], nfold=10, ncross=50, varmod.method="lm")
sum_mod=summary(model_hp)
paste("hp: ",sum_mod$rsq)

model_defense=earth(defense~type1*type2+height_m*weight_kg+is_legendary*experience_growth+generation+base_egg_steps*capture_rate+base_happiness+capture_rate+Evolution,Pokemon[complete.cases(Pokemon),-1], nfold=10, ncross=50, varmod.method="lm")
sum_mod=summary(model_defense)
paste("defense: ",sum_mod$rsq)

model_sp_attack=earth(sp_attack~type1*type2+height_m*weight_kg+is_legendary*experience_growth+generation+base_egg_steps*capture_rate+base_happiness+capture_rate+Evolution,Pokemon[complete.cases(Pokemon),-1], nfold=10, ncross=50, varmod.method="lm")
sum_mod=summary(model_sp_attack)
paste("sp_attack: ",sum_mod$rsq)

model_sp_defense=earth(sp_defense~type1*type2+height_m*weight_kg+is_legendary*experience_growth+generation+base_egg_steps*capture_rate+base_happiness+capture_rate+Evolution,Pokemon[complete.cases(Pokemon),-1], nfold=10, ncross=50, varmod.method="lm")
sum_mod=summary(model_sp_defense)
paste("sp_defense: ",sum_mod$rsq)

model_speed=earth(speed~type1*type2+height_m*weight_kg+is_legendary*experience_growth+generation+base_egg_steps*capture_rate+base_happiness+capture_rate+Evolution,Pokemon[complete.cases(Pokemon),-1], nfold=10, ncross=50, varmod.method="lm")
sum_mod=summary(model_speed)
paste("speed: ",sum_mod$rsq)
```

We've taken HP and will include in in our models:

```{r}
Pokemon_Complete=Pokemon
Pokemon_Complete$Pred_HP=Predict_model_hp$fit
Predict_model_base_order=Predict_model_hp[order(Pokemon_Complete$Pred_HP),]
Pokemon_Complete_order=Pokemon_Complete[order(Pokemon_Complete$Pred_HP),]
Pokemon_Complete_order$index=seq(1:nrow(Pokemon_Complete_order))

ggplot(Pokemon_Complete_order, aes(index, Pred_HP))+
    geom_point()+
    geom_point(aes(y = hp, colour=ifelse(hp<Predict_model_base_order$lwr |
                                            hp>Predict_model_base_order$upr,"orange","red")))+
    geom_line(data=Pokemon_Complete_order)+
    geom_ribbon(aes(ymin=Predict_model_base_order$lwr,ymax=Predict_model_base_order$upr),alpha=0.3) +xlab("Pokemon sorted by Predicted HP")+
  ylab("HP") + ggtitle("Predictions and True Values") + theme(legend.position = "none")+
  theme(plot.title = element_text(hjust = 0.5))

```


```{r}
Pokemon_Complete=Pokemon[complete.cases(Pokemon),]
Pokemon_Complete$Pred_HP<-predict(model_hp)
#View(Pokemon_Complete)

model_attack=earth(attack~type1*type2+height_m*weight_kg+is_legendary*experience_growth+generation+base_egg_steps*capture_rate+base_happiness+capture_rate+Pred_HP,Pokemon_Complete, nfold=10, ncross=50, varmod.method="lm")
sum_mod=summary(model_attack)
paste("attack: ",sum_mod$rsq)

model_defense=earth(defense~type1*type2+height_m*weight_kg+is_legendary*experience_growth+generation+base_egg_steps*capture_rate+base_happiness+capture_rate+Pred_HP,Pokemon_Complete, nfold=10, ncross=50, varmod.method="lm")
sum_mod=summary(model_defense)
paste("defense: ",sum_mod$rsq)

model_sp_attack=earth(sp_attack~type1*type2+height_m*weight_kg+is_legendary*experience_growth+generation+base_egg_steps*capture_rate+base_happiness+capture_rate+Pred_HP,Pokemon_Complete, nfold=10, ncross=50, varmod.method="lm")
sum_mod=summary(model_sp_attack)
paste("sp_attack: ",sum_mod$rsq)

model_sp_defense=earth(sp_defense~type1*type2+height_m*weight_kg+is_legendary*experience_growth+generation+base_egg_steps*capture_rate+base_happiness+capture_rate+Pred_HP,Pokemon_Complete, nfold=10, ncross=50, varmod.method="lm")
sum_mod=summary(model_sp_defense)
paste("sp_defense: ",sum_mod$rsq)

model_speed=earth(speed~type1*type2+height_m*weight_kg+is_legendary*experience_growth+generation+base_egg_steps*capture_rate+base_happiness+capture_rate+Pred_HP,Pokemon_Complete, nfold=10, ncross=50, varmod.method="lm")
sum_mod=summary(model_speed)
paste("speed: ",sum_mod$rsq)
```

Attack is the next highest:

```{r}
Pokemon_Complete$Pred_Attack<-predict(model_attack)

model_defense=earth(defense~type1*type2+height_m*weight_kg+is_legendary*experience_growth+generation+base_egg_steps*capture_rate+base_happiness+capture_rate+Pred_HP+Pred_Attack,Pokemon_Complete)
sum_mod=summary(model_defense)
paste("defense: ",sum_mod$rsq)

model_sp_defense=earth(sp_defense~type1*type2+height_m*weight_kg+is_legendary*experience_growth+generation+base_egg_steps*capture_rate+base_happiness+capture_rate+Pred_HP+Pred_Attack,Pokemon_Complete)
sum_mod=summary(model_sp_defense)
paste("sp_defense: ",sum_mod$rsq)

model_speed=earth(speed~type1*type2+height_m*weight_kg+is_legendary*experience_growth+generation+base_egg_steps*capture_rate+base_happiness+capture_rate+Pred_HP+Pred_Attack,Pokemon_Complete)
sum_mod=summary(model_speed)
paste("speed: ",sum_mod$rsq)
```

Now we'll take special attack:

```{r}
Pokemon_Complete$Pred_SP_Attack<-predict(model_sp_attack)

model_defense=earth(defense~type1*type2+height_m*weight_kg+is_legendary*experience_growth+generation+base_egg_steps*capture_rate+base_happiness+capture_rate+Pred_HP+Pred_Attack+Pred_SP_Attack,Pokemon_Complete)
sum_mod=summary(model_defense)
paste("defense: ",sum_mod$rsq)

model_sp_defense=earth(sp_defense~type1*type2+height_m*weight_kg+is_legendary*experience_growth+generation+base_egg_steps*capture_rate+base_happiness+capture_rate+Pred_HP+Pred_Attack+Pred_SP_Attack,Pokemon_Complete)
sum_mod=summary(model_sp_defense)
paste("sp_defense: ",sum_mod$rsq)

model_speed=earth(speed~type1*type2+height_m*weight_kg+is_legendary*experience_growth+generation+base_egg_steps*capture_rate+base_happiness+capture_rate+Pred_HP+Pred_Attack+Pred_SP_Attack,Pokemon_Complete)
sum_mod=summary(model_speed)
paste("speed: ",sum_mod$rsq)
```

Now we'll take special defense:

```{r}
Pokemon_Complete$Pred_SP_Defense<-predict(model_sp_defense)

model_defense=earth(defense~type1*type2+height_m*weight_kg+is_legendary*experience_growth+generation+base_egg_steps*capture_rate+base_happiness+capture_rate+Pred_HP+Pred_Attack+Pred_SP_Attack+Pred_SP_Defense,Pokemon_Complete)
sum_mod=summary(model_defense)
paste("defense: ",sum_mod$rsq)

model_speed=earth(speed~type1*type2+height_m*weight_kg+is_legendary*experience_growth+generation+base_egg_steps*capture_rate+base_happiness+capture_rate+Pred_HP+Pred_Attack+Pred_SP_Attack+Pred_SP_Defense,Pokemon_Complete)
sum_mod=summary(model_speed)
paste("speed: ",sum_mod$rsq)
```

Finally defense to find speed:

```{r}
Pokemon_Complete$Pred_defense<-predict(model_defense)

model_speed=earth(speed~type1*type2+height_m*weight_kg+is_legendary*experience_growth+generation+base_egg_steps*capture_rate+base_happiness+capture_rate+Pred_HP+Pred_Attack+Pred_SP_Attack+Pred_SP_Defense+Pred_defense,Pokemon_Complete, nfold=10, ncross=30, varmod.method="lm")
Predict_model_base=predict(model_speed,interval="pint")

sum_mod=summary(model_speed)
paste("speed: ",sum_mod$rsq)

model_base=earth(base_total~type1*type2+height_m*weight_kg+is_legendary*experience_growth+generation+base_egg_steps*capture_rate+base_happiness+capture_rate+Pred_HP+Pred_Attack+Pred_SP_Attack+Pred_SP_Defense+Pred_defense,Pokemon_Complete, nfold=10, ncross=30, varmod.method="lm", keepxy=TRUE)
Predict_model_base=predict(model_base,interval="pint")
sum_mod=summary(model_base)
paste("base: ",sum_mod$rsq)
mean(model_base$cv.rsq.tab[model_base$cv.rsq.tab[,1]>0])
sum(model_base$cv.rsq.tab[,1]<0)/length(model_base$cv.rsq.tab[,1])
```

Note that about 5% of our total $R^2$ values are erronous.

```{r}
Pokemon_Complete$Pred_base<-predict(model_base)
Pokemon_Complete$Pred_speed<-Pokemon_Complete$Pred_base-Pokemon_Complete$Pred_HP-Pokemon_Complete$Pred_Attack-Pokemon_Complete$Pred_SP_Attack-Pokemon_Complete$Pred_SP_Defense-Pokemon_Complete$Pred_defense
```

```{r}
#Pokemon_Complete_order=Pokemon_Complete[order(Pokemon_Complete$Pred_base),]
#Pokemon_Complete_order$index=seq(1:nrow(Pokemon_Complete_order))
#plot(Pokemon_Complete_order$Pred_base)
#points(Pokemon_Complete_order$base_total,col="red")

#Pokemon_Complete_order=Pokemon_Complete[order(Pokemon_Complete$Pred_BaseOnly),]
#plot(Pokemon_Complete_order$Pred_BaseOnly)
#points(Pokemon_Complete_order$base_total,col="red")




#Predict_model_base_only
Pokemon_Complete_order=Pokemon_Complete[order(Pokemon_Complete$Pred_Base_only),]
Pokemon_Complete_order$index=seq(1:nrow(Pokemon_Complete_order))
Predict_model_base_order=Predict_model_base_only[order(Predict_model_base_only$fit),]

ggplot(Pokemon_Complete_order, aes(index, Pred_Base_only))+
    geom_point()+
    geom_point(aes(y = base_total, colour = "red"))+
    geom_line(data=Pokemon_Complete_order)+
    geom_ribbon(aes(ymin=Predict_model_base_order$lwr,ymax=Predict_model_base_order$upr),alpha=0.3)

Pokemon_Complete_order=Pokemon_Complete[order(Pokemon_Complete$Pred_base),]
Pokemon_Complete_order$index=seq(1:nrow(Pokemon_Complete_order))
Predict_model_base_order=Predict_model_base[order(Predict_model_base$fit),]

ggplot(Pokemon_Complete_order, aes(index, Pred_base))+
    geom_point()+
    geom_point(aes(y = base_total, colour=ifelse(base_total<Predict_model_base_order$lwr |
                                            base_total>Predict_model_base_order$upr,"orange","red")))+
    geom_line(data=Pokemon_Complete_order)+
    geom_ribbon(aes(ymin=Predict_model_base_order$lwr,ymax=Predict_model_base_order$upr),alpha=0.3) +xlab("Pokemon sorted by Predicted total")+
  ylab("Total Stats") + ggtitle("Predictions and True Values") + theme(legend.position = "none")+
  theme(plot.title = element_text(hjust = 0.5))

#Predict_model_base_only$fit[1:10]
#Predict_model_base$fit[1:10]

```

```{r}
Pokemon_Complete_order=Pokemon_Complete[order(Pokemon_Complete$Pred_speed),]
Pokemon_Complete_order$index=seq(1:nrow(Pokemon_Complete_order))

ggplot(Pokemon_Complete_order, aes(index, Pred_speed))+
    geom_point(aes(y = speed))+
    geom_line(data=Pokemon_Complete_order)+
  ylab("Total Stats") + ggtitle("Predictions and True Values")+
  theme(plot.title = element_text(hjust = 0.5))
```

